---
title: "NPPES"
author: "Mohamed Megahed"
date: "2/18/2020"
output: html_document
---

# first as our dataset size is 7 GB, we can import it using he following
```{r setup, include=FALSE}
N <-read.csv(file="npidata_pfile_20050523-20200209.csv",
              header= TRUE,
              sep=",",
              nrows = 200000)
ncol(N)
nrow(N)
```

### Data Pre-processing

## checking if there is any duplicated records
```{r}
Duplicated_RECORDS <- N[N$NPI==N$NPI[duplicated(N$NPI)],]
Duplicated_RECORDS
```

** checking if there any missing data using "Amelia package"
```{r}
# Check if there are any missing values:
anyNA(N)
sum(is.na(N))
mean(is.na(N))
# the percent of missing data values is 24.08 %

# Check percentages and counts of missing values in columns:
col_miss_per <- colMeans(is.na(N))*100
col_miss <- as.data.frame(col_miss_per)
col_mi <- col_miss[order(col_miss_per,decreasing = TRUE),]
col_mi
colSums(is.na(N))
```

# ploting the missing and observed data values
```{r, fig.height=7, fig.width=15}
# A visual take on the missing values might be helpful: the Amelia package has a special plotting function missmap() that will plot the dataset and highlight missing values:
library(Amelia)
missmap(N, main = "Missing values vs observed")
```

### taking random samples
```{r}
set.seed(1000)
R1 <- N[sample(nrow(N), 50000),]
R2 <- N[sample(nrow(N), 50000),]
```

# compare the 2 daasets and gettig the rows that are not exists in both datasets and the rows that are exists in both datasets
```{r}
library("sqldf")
R1NotInR2 <- sqldf('SELECT * FROM R1 EXCEPT SELECT * FROM R2')
R1NotInR2
#And the rows which are in both data frames:
R1InR2 <- sqldf('SELECT * FROM R1 INTERSECT SELECT * FROM R2')
R1InR2
```

```{r}
library(compareDF)
R_compare = compare_df(R1[1:100,], R2[1:100,], "NPI")
head(R_compare$comparison_df)
```

# look for similarity 
```{r}
UNI <- unique(N[c("NPI","Healthcare.Provider.Primary.Taxonomy.Switch_1")])
UNI

UNI2 <- unique(N[c("NPI","Provider.Business.Practice.Location.Address.City.Name","Entity.Type.Code")])
UNI2
```

# Unique
```{r}
unique(N)
```

```{r}
# Using subset:
missing<-subset(R1, !(NPI %in% R2$NPI))
missing
```

# you want to see if each value of R1$NPI (for instance) matches the entire values of R2$Provider.Organization.Name..Legal.Business.Name. (for instance).
```{r}
library(tidyverse)
s <- bind_cols(R2, map_dfc(R1$NPI, ~ transmute(R2, new_col = R1$NPI[.] == R2$Provider.Organization.Name..Legal.Business.Name.)))
head(s)
```