---
title: "Case Study 3"
author: "Mohamed Megahed"
date: "4/19/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---
```{r, echo=FALSE}
knitr::opts_chunk$set(error = TRUE)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Class Practice - 8
```{r}
library(ggplot2)
library(ggforce)
library(gghighlight)
```

__1. plot the following graph and label as shown:__
_ggplot(economics, aes(date, unemploy)) +  geom_line() + labs(title=caption,  y="Unemployment",  x="Year")_ 
```{r, fig.height=5, fig.width=11}
ggplot(economics, aes(date, unemploy)) +  geom_line() + labs(title="Umeployment rates",  y="Unemployment",  x="Year",caption = "Umemployment rates")
```

__2 Plot following program to annotate the graph__
```{r, fig.height=5, fig.width=11}

p <- ggplot(mpg, aes(displ, cty)) +
geom_point(data = filter(mpg, manufacturer == "audi"),colour = "red",
size = 3
) +geom_point()

p +
annotate(
geom = "curve", x = 3, y = 35, xend = 2.8, yend = 20,
curvature = .3, arrow = arrow(length = unit(2, "mm"))
) +
annotate(geom = "text", x = 3.1, y = 35, label = "Audi", hjust = "left")

ggplot(mpg, aes(displ, hwy)) +
  geom_point(colour = "red") + 
  ggforce::geom_mark_ellipse(aes(label = cyl, group = cyl, color = cyl))

ggplot(mpg, aes(displ, cty, 
            colour = factor(drv))) +
  geom_point() + 
  gghighlight::gghighlight() + 
  facet_wrap(vars(drv))
```

__3 Try following plot__
```{r, fig.height=5, fig.width=11}
#Plot 1:
ggplot(mpg, aes( displ,cty)) + 
  geom_point() + 
  geom_smooth()

#Plot 2:
ggplot(mpg, aes(cty,displ)) + 
  geom_point() + 
  geom_smooth()

#Plot 3:
ggplot(mpg, aes(displ, cty)) + 
  geom_point() + 
  geom_smooth() + 
  coord_flip()

```

__4. Using "Annotation" discussed earlier, highlight car type "Honda"__
```{r, fig.height=5, fig.width=11}
h <- ggplot(mpg, aes(displ, cty)) +
  geom_point(
    data = filter(mpg, manufacturer == "honda"), 
    colour = "green",
    size = 3
           ) + geom_point() + theme_minimal()

h + annotate(
    geom = "curve", x = 4, y = 27, xend = 2.05, yend = 21.5, 
    curvature = .3, arrow = arrow(length = unit(3, "mm"))
    ) + annotate(geom = "text", x = 4.02, y = 27, label = "Honda", hjust = "left",colour = "red")
```

__5. Using ggforce() to highlight the "Class" of cars on "City" mileage based on engine (displ)__
```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(displ, cty)) +
  geom_point(size = 1, colour = "blue") + theme_minimal() +
  ggforce::geom_mark_ellipse(aes(label = class, group = class,color = class))
```

__6. Using gghighlight() cluster "transmission" type__
```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(displ, cty, 
            colour =factor(trans))) +
  geom_point(size = 1) + theme_minimal() +
  gghighlight::gghighlight() + 
  facet_wrap(vars(trans))
```

# Class Practice - 9

### Building layers:

_1. Execute the following commands to build the first layer:_
_p<-ggplot(mpg, aes(displ, hwy))_
```{r, fig.height=5, fig.width=11}
p<-ggplot(mpg, aes(displ, hwy))
p
```
__2 Add aesthetics to the plot:__

```{r, fig.height=5, fig.width=11}
p + layer(
  mapping = NULL, 
  data = NULL,
  geom = "point", 
  stat = "identity",
  position = "identity"
)
```

__3. Can you change "colour" of points?__
```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(displ, hwy))+ layer(
  mapping = NULL, 
  data = NULL,
  geom = "point", 
  stat = "identity",
  position = "identity",
)+ geom_point(size = 1, colour = "red")
```

__4. Fit the line to the data points by executing following ggplot() program:__

```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(displ, hwy, colour = class)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  theme(legend.position = "none")
```

__5. Add two methods in the geom_smooth__
```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = class)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(legend.position = "none")
```