---
title: "Case Study 2"
author: "Mohamed Megahed"
date: "4/19/2020"
output:
  word_document: default
  html_document: default
---
# Class Practice - 5 (Univariate Analysis)
```{r}
library(ggplot2)
```
__1. Plot the histogram using ggplot()__

_ggplot(cars, aes(hwy)) + geom_histogram()_

```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(hwy)) + geom_histogram(fill = "red")+ labs(x = "Highway", title = "Miles per gallon") +
 theme_minimal()
```

__2. Plot the distribution of the variables using geom_freqpoly()__
_ggplot(cars, aes(hwy)) + geeom_freqpoly()_
```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(hwy)) + geom_freqpoly(size = 0.9,colour = "blue") + labs(x = "Highway", title = "Frequency Plot - Highway") + theme_minimal()
```

__3.	Plot box plot using geom_boxplot()__
```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(class,cty)) + geom_boxplot(size = 0.5,colour = "blue") + labs(x = "Class",y= "City (Miles per gallon)", title = "Boxplot - Class vs City(mpg)") + theme_minimal()
```

__4	What is the purpose of Histograms and Density plot?__

_Histogram Plot: we use it to visualize the type of distribution usually for continous data. as it is a great way to get started exploring a single variable. A histogram divides the variable into bins, counts the data points in each bin, and shows the bins on the x-axis and the counts on the y-axis.

Density plots are usually a much more effective way to view the distribution of a variable

__5 Name another Univariate plot?__
_geom_density()_
```{r}
ggplot(mpg, aes(hwy)) + geom_density(size = 0.8,colour = "#0c4c8a") + labs(x = "Highway(Miles per gallon)", title = "Density plot- Highway(miles per gallon)") + theme_minimal()
```

__6. Plot the following using ggplot() + facet_wrap()__

_ggplot(mpg, aes(cty, fill = drv)) +   geom_histogram(binwidth = 0.5) +  facet_wrap(~drv, ncol = 1)_

```{r}
ggplot(mpg, aes(cty, fill = drv)) +   geom_histogram(binwidth = 0.5) +  facet_wrap(~drv, ncol = 1) + labs(x = "City(Miles per gallon)", title = "Histogram - Drive") + theme_minimal()
```

__7.	Plot the following graph:__

```{r, fig.height=5, fig.width=11}
ggplot(mpg, aes(displ, colour = class)) + geom_freqpoly(binwidth = 0.5) + labs(x = "City(Miles per gallon)", title = "Class - Frequency ploygon Plot")
```

# Class Practice - 6 

### Time Series: 

__1.	Use the 'economics' dataset and Plot the following using geom_line():__
_ggplot(economics, aes(date, pop)) + geom_line()_
```{r, fig.height=5, fig.width=11}
ggplot(economics, aes(date, pop)) + geom_line()
```

__2.Plot the unemployment growth rate over a period of time:__
_ggplot(economics, aes(date, unemploy)) +  geom_line()_

```{r, fig.height=5, fig.width=11}
ggplot(economics, aes(date, unemploy)) +  geom_line()
```

__3 Plot the graph to show how long people were unemployed?__
```{r, fig.height=5, fig.width=11}
ggplot(economics, aes(unemploy / pop, uempmed)) + geom_line(colour = "#0c4c8a") + ggtitle("Time Series Plot of people unemployed duration") + theme_minimal()
```
__4	Plot the below graph:__
_(HINT: Use POSIXlt() ; ggplot() +  geom_path() + geom_point()_

```{r, fig.height=5, fig.width=11}
year <- function(x) as.POSIXlt(x)$year + 1900
ggplot(economics, aes(unemploy / pop, uempmed)) + geom_path() + geom_point(aes(colour = year(date)))+ ggtitle("Time series plot model")
```

# Class Practice - 7

__1.Explore the distribution of the 'carat' variable in the diamonds dataset. What binwidth reveals the most interesting patterns?__
```{r, fig.height=5, fig.width=11}
ggplot(diamonds,aes(carat))+geom_histogram(binwidth=0.5,fill = "blue")+theme_minimal()
ggplot(diamonds,aes(carat))+geom_histogram(binwidth=0.1,fill = "blue")+theme_minimal()

# Using different values of bins
ggplot(diamonds,aes(carat))+geom_histogram(bins=50,fill = "red")+theme_minimal()
ggplot(diamonds,aes(carat))+geom_histogram(bins=100,fill = "red")+theme_minimal()
ggplot(diamonds,aes(carat))+geom_histogram(bins=150,fill = "red")+theme_minimal()
ggplot(diamonds,aes(carat))+geom_histogram(bins=250,fill = "red")+theme_minimal()
```

__2.	Explore the distribution of the 'price' variable in the diamonds data. How does the distribution vary by cut?__
```{r, fig.height=5, fig.width=11}
ggplot(diamonds,aes(price,fill=cut))+ geom_density(alpha=0.3) + ggtitle("Density Distribution - Price Vs Cut")+theme_minimal()
```